The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2012-09-28 script/de/script-de.tex
CT Script: Many minor corrections and updates (e.g. 4 levels of titles)
Many small corrections and enhancements in the CT Script

\floatstyle{ruled}
\newfloat{sagecode}{!ht}{loc}[chapter]
\floatname{sagecode}{Sage-Beispiel}
\newfloat{cryptoprocedure}{!ht}{lop}[chapter]
\floatname{cryptoprocedure}{Krypto-Verfahren}

@@

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begingroup
% remove extra vertical space from \listoffigure and \listoftables so that 
% it look similar to \listof{sagecode}
\setlength{\parskip}{0pt}
\clearpage\phantomsection
\addcontentsline{toc}{chapter}{\listfigurename}
\listoffigures

@@

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Verzeichnis der Krypto-Verfahren mit Pseudocode}
\listof{cryptoprocedure}{Verzeichnis der Krypto-Verfahren mit Pseudocode}
\endgroup

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Verzeichnis der Sage-Programmbeispiele}
\listof{sagecode}{Verzeichnis der Sage-Programmbeispiele}
\label{sc:List-of-Sage-Code-Examples}

\vspace{30pt}
\begin{itemize}
  \item Alle Beispiel wurden mit Sage v 4.7 (Release Date 2011-05-23) getestet.
  \item Der Quellcode der Sage-Beispiele in diesem Skript wird in Form von
        Sage-Programmdatei"-en mit dem CrypTool 1-Setup-Programm ausgeliefert.
        Nach der Installation von CrypTool 1 \index{CrypTool 1} finden Sie
        diese im Unterverzeichnis \verb#sage# innerhalb des
        CrypTool 1-Verzeichnisses:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}   % xxxxxxxxxxxxxx Weitere Beispiele ergänzen !

>>>>

\floatstyle{ruled}

\newfloat{ctsquotes}{!ht}{loc}[chapter]   % ctsQuotes-from-Books
\floatname{ctsquotes}{Verzeichnis der Zitate}

\newfloat{sagecode}{!ht}{loc}[chapter]
\floatname{sagecode}{Sage-Beispiel}

\newfloat{cryptoprocedure}{!ht}{lop}[chapter]
\floatname{cryptoprocedure}{Krypto-Verfahren}

@@

% \setcounter{tocdepth}{¿Anzahl der Ebenen¿} Default ist 3
% \setcounter{secnumdepth}{¿Nummerierungstiefe¿}
\setcounter{tocdepth}{4}    % Das Inhaltsverzeichnis wird vier Ebenen tief
\setcounter{secnumdepth}{4} % Nummerierung auch bei subsubsection

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begingroup
  % remove extra vertical space from \listoffigure and \listoftables so that 
  % it look similar to \listof{sagecode}
  \setlength{\parskip}{0pt}

  \clearpage\phantomsection
  \addcontentsline{toc}{chapter}{\listfigurename}
  \listoffigures

@@

  \clearpage\phantomsection
  \addcontentsline{toc}{chapter}{Verzeichnis der Krypto-Verfahren mit Pseudocode}
  \listof{cryptoprocedure}{Verzeichnis der Krypto-Verfahren mit Pseudocode}

\endgroup

% Wenn das aufgerufen, fehlen die Programmbeispiele unter "Verzeichnis der Sage-Bsp"
% und unter zitate-Verzeichnis steht eine lange undefinierte Liste.
% TODOTODO (Joerg) klären
%\clearpage\phantomsection
%\addcontentsline{toc}{chapter}{Verzeichnis der Zitate}
%\listof{ctsquotes}{Verzeichnis der Zitate}
%\index{Zitate}
%\label{sc:List-of-Quotes}

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Verzeichnis der Sage-Programmbeispiele}
\listof{sagecode}{Verzeichnis der Sage-Programmbeispiele}
\index{Sage!Programmbeispiele}
\label{sc:List-of-Sage-Code-Examples}
% \hypertarget{sc:List-of-Sage-Code-Examples}{}

\vspace{30pt}
\begin{itemize}
  \item Der Quellcode der Sage-Beispiele in diesem Skript wird in Form von
        Sage-Programmdatei"-en mit dem CrypTool 1-Setup-Programm ausgeliefert.
        Die Beispiele in einem Kapitel sind jeweils in einer Datei zusammen gefasst.\\
        Nach der Installation von CrypTool 1 \index{CrypTool 1} finden Sie
        die Sage-Beispiele im Unterverzeichnis \verb#sage# innerhalb des
        CrypTool 1-Verzeichnisses:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
  \item Alle Beispiel wurden mit der Sage-Version 5.3 (Release Date 2012-09-08) getestet.
\end{itemize}   % xxxxxxxxxxxxxx Weitere Beispiele ergänzen !

### 2012-09-28 script/en/chapters/aboutcryptool.tex
Many small corrections and enhancements in the CT Script

\parskip 4pt
%\vskip +12 pt
\noindent In this {\em script accompanying the CrypTool programs} \index{CrypTool}
you will find predominantly mathematically oriented information on using
cryptographic procedures. Also included are many sample code pieces written in the
computer algebra system {\bf Sage}\index{Sage} (see appendix~\ref{s:appendix-using-sage}).
The main chapters have been written by various {\bf authors}
(see appendix~\ref{s:appendix-authors}) %\hyperlink{appendix-authors}{authors}
and are therefore independent from one another. At the end of most chapters
you will find references and web links.

@@

Whereas the CrypTool \textit{e-learning programs}\index{e-learning} motivate
and teach you how to use cryptography in practice, the \textit{script} provides
those interested in the subject with a deeper understanding of the mathematical
algorithms used -- trying to do it in an instructive way.
If you are already a little bit familiar with this field of knowledge you can
gain a fast overview about the functions delivered by the different CrypTool
variants looking at the {\bf menu trees} (see appendix \ref{s:appendix-menu-overview-CT1} etc.).

>>>>

\parskip 4pt
%\vskip +12 pt
\noindent In this {\em script accompanying the CrypTool programs} \index{CrypTool}
you will find predominantly mathematically oriented information on using
cryptographic procedures. Also included are many sample code pieces written in the
computer algebra system {\bf Sage}\index{Sage} (see appendix~\ref{s:appendix-using-sage}).
The main chapters have been written by various {\bf authors}
(see appendix~\ref{s:appendix-authors}) %\hyperlink{appendix-authors}{authors}
and are therefore independent from one another. At the end of most chapters
you will find references and web links.
The sections have been enriched with many footnotes. Within the footnotes you can see
where the described functions can be found in the different CrypTool versions.

@@

Whereas the CrypTool \textit{e-learning programs}\index{e-learning} motivate
and teach you how to use cryptography in practice, the \textit{script} provides
those interested in the subject with a deeper understanding of the mathematical
algorithms used -- trying to do it in an instructive way.
If you are already a little bit familiar with this field of knowledge you can
gain a fast overview about the functions delivered by the different CrypTool
variants looking at the {\bf menu trees} (see appendix \ref{s:appendix-menu-overview-CT1}, \ref{s:appendix-template-overview-CT2} and \ref{s:appendix-function-overview-JCT}).

### 2012-09-28 script/en/chapters/introduction.tex
Many small corrections and enhancements in the CT Script

Starting in the year 2000 this script became part of the CrypTool 1
(CT1)\index{CrypTool 1.x} package. It is designed to accompany the
program CrypTool by explaining some mathematical topics in more detail, 
but still in a way which is easy to understand.

@@

This edition of the CT script again updated some topics:
%This edition completely updated the TeX sources of the document, and of course
%the content of the script was corrected, amended and updated with some topics, e.g.:
\vspace{-5pt}
\begin{itemize}
  \item the definitions of the strength of security functions
        (Kap. \ref{cm_Section_Security_Definitions}),
  \item the largest prime numbers (chap. \ref{search_for_very_big_primes}),
        new factorization records (chap. \ref{RSA-768}),
%  \item progress in cryptanalysis of AES 
%        (chap. \ref{NeueAES-Analyse}) and
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played a major role (see appendix \ref{s:appendix-movies});
        and where primes are used as hangers  
        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}).
  \item Newly added are the functions overviews 
        \hyperlink{template-overview-CT2}{for CT2} and
        \hyperlink{function-overview-JCT}{(for JCT},
        further Sage scripts for cryptography and the appendix
        \ref{s:appendix-using-sage} about using the computer algebra system
        Sage. Sage becomes more and more the standard open-source CAS system.
        % Nguyen and Massierer (see also chap.~\ref{ec:Sage_Massierer}).
\end{itemize}

@@

I hope that many readers have fun with this script and that they get 
out of it more interest and greater understanding of this modern but 
also very ancient topic.
\\
\\
% [1.5\baselineskip]
% \enlargethispage*{2\baselineskip}
% \nopagebreak
Bernhard Esslinger
\\
\\
% [\baselineskip]
Frankfurt (Germany), July 2012

@@

CrypTool 1\index{CrypTool 1} and the successor versions
CrypTool 2 (CT2)\index{CrypTool 2.0} and JCrypTool (JCT)\index{JCrypTool 1.0}
are used world-wide for training in companies and teaching at schools and
universities.
\par \vskip + 15pt

>>>>

Starting in the year 2000 this script became part of the CrypTool 1
(CT1)\index{CrypTool 1} package. The script's goal is to explain some mathematical
topics in more detail, but still in a way which is easy to understand.

@@

%This edition of the CT script again updated some topics:
This edition completely updated the TeX sources of the document, and of course
the content of the script was amended, corrected, and updated with some topics, e.g.:
\vspace{-5pt}
\begin{itemize}
  \item the definitions of the strength of security functions
        (chap. \ref{cm_Section_Security_Definitions}),
  \item the largest prime numbers (chap. \ref{search_for_very_big_primes}),
        new factorization records (chap. \ref{RSA-768}),
%  \item progress in cryptanalysis of AES 
%        (chap. \ref{NeueAES-Analyse}) and
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played a major role (see appendix \ref{s:appendix-movies});
        and where primes are used as hangers  
        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}),
  \item the overviews of all functions in
        \hyperlink{template-overview-CT2}{for CT2} and
        \hyperlink{function-overview-JCT}{for JCT},
  \item further Sage scripts for cryptography, and the appendix
        \ref{s:appendix-using-sage} about using the computer algebra system
        Sage. Sage becomes more and more the standard open-source CAS system,
        % Nguyen and Massierer (see also chap.~\ref{ec:Sage_Massierer}).
  \item the section about RSA fixed points (see \ref{l:NumberTheory_Sage_Number-of-RSA-FixedPoints}), and
%  \item the chapter about homomorphic encryption.
\end{itemize}

@@

I hope that many readers have fun with this script and that they get 
out of it more interest and greater understanding of this modern but 
also very ancient topic.
\\
\\
% [1.5\baselineskip]
% \enlargethispage*{2\baselineskip}
% \nopagebreak
Bernhard Esslinger
\\
\\
% [\baselineskip]
Frankfurt (Germany), September 2012

@@

CrypTool 1\index{CrypTool 1} and the successor versions
CrypTool 2 (CT2)\index{CrypTool 2} and JCrypTool (JCT)\index{JCrypTool}
are used world-wide for training in companies and teaching at schools and
universities.
\par \vskip + 15pt

### 2012-09-28 script/en/chapters/menus.tex
Many small corrections and enhancements in the CT Script

This appendix contains at the following page the complete menu tree of
CrypTool\index{CrypTool} version 1.4.31\footnote{%
  Since 2010, changes for the stable CrypTool 1.x (CT1)\index{CrypTool 1.x}
  focus mainly on bugfixes. All new developments go into the two successor
  versions CrypTool 2 (CT2)\index{CrypTool 2.0} and
  JCrypTool (JCT)\index{JCrypTool}:\\
  - Web site CT2: \url{http://www.cryptool.org/en/ct2-documentation-en} \\
  - Web site JCT: \url{http://www.cryptool.org/en/jct-volunteer-en} \\
  These successor versions are currently (July 2012) still beta versions;
  but they are close to their first release and they are already stable
  enough to be used by end-users.\\
}.

>>>>

This appendix contains at the following page the complete menu tree of
CrypTool\index{CrypTool} version 1.4.31\footnote{%
  Since 2010, changes for the stable CrypTool 1 (CT1)\index{CrypTool 1}
  focus mainly on bugfixes. All new developments go into the two successor
  versions CrypTool 2 (CT2)\index{CrypTool 2} and
  JCrypTool (JCT)\index{JCrypTool}:\\
  - Web site CT2: \url{http://www.cryptool.org/en/ct2-documentation-en} \\
  - Web site JCT: \url{http://www.cryptool.org/en/jct-volunteer-en} \\
  These successor versions are currently (July 2012) still beta versions;
  but they are close to their first release and they are already stable
  enough to be used by end-users.\\
}.

### 2012-09-28 script/en/chapters/movies.tex
Many small corrections and enhancements in the CT Script

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{appendix-movies}{}%neu
\section{Movies and Fictional Literature with Relation to Cryptography, Books for Kids 
with Simple Ciphers}
\label{s:appendix-movies}
% {\bf movies and literature related to cryptography}(see appendix \ref{s:appendix-movies})
\index{Movies}
\index{Literature}

@@

\item[\textrm{[Suarez2009]}] \index{Suarez 2009}
    Daniel Suarez, \\
    {\em Daemon}, Penguin Books, 2009. \\
    This is considered as one of the most exciting books during the last few years -- its a
    near-science-fiction thriller combining developments in the real world and possibilities
    coming from current research like from the Google X Lab (augmented reality
    head-mounted displays (HMD) like Google glass, self-driving cars, 3-D printers, ...)
    to a plausible story.

@@

@@

The following list contains movies and ``kid books''.
The kid books here contain collections of simpler cryptographic encryption
methods, prepared in a didactic and exciting manner
(please send us similar English kid books and kid movies, because at the moment
our list contains only German kid books):

>>>>

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{appendix-movies}{}%neu
\section{Movies and Fictional Literature with Relation to Cryptography}
\label{s:appendix-movies}
% {\bf movies and literature related to cryptography}(see appendix \ref{s:appendix-movies})
\index{Movies}
\index{Literature}

@@

\item[\textrm{[Suarez2009]}] \index{Suarez 2009}
    Daniel Suarez, \\
    {\em Daemon}, Penguin Books, 2009, 632 pages. \\
    This is considered as one of the most exciting books during the last few years -- its a
    near-science-fiction thriller combining developments in the real world and possibilities
    coming from current research like from the Google X Lab (augmented reality
    head-mounted displays (HMD) like Google glass, self-driving cars, 3-D printers, ...)
    to a plausible story.

@@

\item[\textrm{[Suarez2010]}] \index{Suarez 2010}
    Daniel Suarez, \\
    {\em Freedom (TM)}, Penguin Books, 2010, 486 pages. \\
    ``The propulsive, shockingly plausible sequel to the bestseller Daemon'' (see above).
    Freedom (TM) (Daemon \#2) patches a number of holes the writer left in the first book.
    The prose is tighter, the descriptions more direct, the characters are fleshed out.
    Having laid all the groundwork in ``Daemon'', Suarez uses most of this book to use this
    foundation in order to explore a new concept of social organization based on empowering
    information technology and the reasoning why and how the battle runs between the old
    potentates and the Daemon society, which also evolves further during the story.
    Cryptography is a natural part of modern technology and modern warfare as described
    in this book.
    The new society emerging in "Freedom (TM)" is based on the darknet, an alternative to
    the internet using fast wireless meshes in order to increase the durability and
    availability of the network. Its not far away from the parallel usage of modern
    technology integrated into our modern lives as a virtual world overlaying our real
    world.\\

\item[\textrm{[Burger2011]}] \index{Burger 2011}
    Wolfgang Burger, \\
    {\em The fifth murderer (original title: Der f\"unfte M\"order)}, Piper, 2011. \\
    Location \& time of the story: Germany / Heidelberg, 1990 - 2009. 
    Episode 7 of the Alexander-Gerlach series.
    Inspector Alexander Gerlach became almost a victim of a bomb blast when the
    sport utility vehicle (SUV) of a Bulgarian panderer exploded.
    Gerlach starts investigating because he wants to prevent a gang warfare, but then
    his bosses call him off.
    When the journalist Machatschek supports Gerlach, he communicates with him only via
    Skype using an add-on encryption program which he believes is the most secure in the
    world.\\

@@

The following list contains movies and ``kid books''.
The kid books contain both stories, and collections of simple ciphers, prepared
in a didactic and exciting manner
(please send us similar English kid books and kid movies, because at the moment
our list contains mostly German kid books):
